Evolution of precopulatory and postcopulatory inbreeding behaviour

### 2016-09-06 ms_pre_post.Rmd
Fix to only \beta = 3 in the manuscript, and tweaks to the main figures up front

To analyse our model, we track mean values of alleles underlying tendency for polyandry ($P_{a}$), inbreeding mating strategy ($M_{a}$), and inbreeding fertilisation strategy ($F_{a}$) in a focal population. We present evolutionary dynamics of these mean allele values across different combinations of inbreeding depression slopes and costs to test our two focal hypotheses. The combinations of costs and inbreeding depression slopes that we consider include all eight possible permutations of values of costs of tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$) equal to either $0$ or $0.02$ [@Duthie observed $0.02$ to be strong cost of tendency for polyandry], and for values of $\beta = \{1, 3\}$ ($16$ total combinations). In all of these simulations, we record mean values of $P_{a}$, $M_{a}$, $F_{a}$ in each generation. Each of the 16 possible cost combinations is replicated $40$ times, and grand mean values and standard deviation of means is calculated in each generation across replicate simulations. These analyses allowed us to infer how allele values change over generations in response to costs, but also in response to the changing values of other alleles and therefore potentially interacting phenotypes (e.g., selection on inbreeding fertilisation strategy might be weak if polyandry is rare).

To test our hypothesis that selection for evolution of polyandry is driven by female's ability to bias fertilisation toward inbreeding avoidance through post-copulatory mechanisms, we compare simulations in which all three phenotypes including tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$) evolve, with an identical set of simulations in which $F_{p}$ cannot evolve. In simulations without evolving $F_{p}$, we therefore include all 16 combinations of $c_{P}=\{0,0.02\}$, $c_{M}=\{0,0.02\}$, $c_{F}=\{0,0.02\}$, and $\beta = \{1,3\}$ replicated 40 times each. To stop $F_{p}$ from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes resulting from $F_{a}$ cause random fertilisation with respect to kinship. Hence, while mutation and drift can still occur for $F_{a}$ alleles, these alleles have no phenotypic effect and polyandry cannot cause polyandrous females to benefit through an increased ability to bias fertilisation toward inbreeding avoidance (under such conditions, there can also never be a cost of $F_{p}$, meaning that $c_{F}=0$). By comparing evolution of mean $P_{p}$ values over generations when $F_{p}$ also evolves versus when $F_{p}=0$ (random fertilisation), we test whether or not evolution of inbreeding fertilisation strategy increases selection for evolution of polyandry.

>>>>

To analyse our model, we track mean values of alleles underlying tendency for polyandry ($P_{a}$), inbreeding mating strategy ($M_{a}$), and inbreeding fertilisation strategy ($F_{a}$) in a focal population. We present evolutionary dynamics of these mean allele values across different combinations of inbreeding depression slopes and costs to test our two focal hypotheses. The combinations of costs and inbreeding depression slopes that we consider include all eight possible permutations of values of costs of tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$) equal to either $0$ or $0.02$ [@Duthie observed $0.02$ to be strong cost of tendency for polyandry], and for values of $\beta = 3$ ($8$ total combinations). In all of these simulations, we record mean values of $P_{a}$, $M_{a}$, $F_{a}$ in each generation. Each of the 16 possible cost combinations is replicated $40$ times, and grand mean values and standard deviation of means is calculated in each generation across replicate simulations. These analyses allowed us to infer how allele values change over generations in response to costs, but also in response to the changing values of other alleles and therefore potentially interacting phenotypes (e.g., selection on inbreeding fertilisation strategy might be weak if polyandry is rare).

To test our hypothesis that selection for evolution of polyandry is driven by female's ability to bias fertilisation toward inbreeding avoidance through post-copulatory mechanisms, we compare simulations in which all three phenotypes including tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$) evolve, with an identical set of simulations in which $F_{p}$ cannot evolve. In simulations without evolving $F_{p}$, we therefore include all 16 combinations of $c_{P}=\{0,0.02\}$, $c_{M}=\{0,0.02\}$, $c_{F}=\{0,0.02\}$, and $\beta = 3$ replicated 40 times each. To stop $F_{p}$ from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes resulting from $F_{a}$ cause random fertilisation with respect to kinship. Hence, while mutation and drift can still occur for $F_{a}$ alleles, these alleles have no phenotypic effect and polyandry cannot cause polyandrous females to benefit through an increased ability to bias fertilisation toward inbreeding avoidance (under such conditions, there can also never be a cost of $F_{p}$, meaning that $c_{F}=0$). By comparing evolution of mean $P_{p}$ values over generations when $F_{p}$ also evolves versus when $F_{p}=0$ (random fertilisation), we test whether or not evolution of inbreeding fertilisation strategy increases selection for evolution of polyandry.
